The development of children's ability to track and predict turn structure in conversation

### 2015-12-11 paper/CasillasFrank-Anticip.tex
a few last paper changes for the resubmission

Interestingly, adults and children both were strongly affected by transition type, in that they made more anticipatory switches after hearing questions, compared to non-questions. Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}). Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari12, de-ruiter2006}. The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to upcoming responses.

@@

However, these results do not support the idea that lexical information is sufficient for children, as has been proposed for adults previously \citealp{de-ruiter2006} and replicated in the present study. On the contrary, children showed significant gains in the \textit{normal} speech condition, but \textit{not} in the \textit{words only} condition. Notably, the \textit{normal} speech condition, in addition to having both lexical and prosodic information, is also the one most likely to occur in our participants' daily lives (compared to muffled or robotic speech) and therefore may have an extra advantage of familiarity over the other conditions.

@@

Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \cite{magyari2012, de-ruiter2006} and replicated in the current study for adult participants: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal. Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues. If this effect arose in our data because children do not make predictions based on phonetically manipulated speech in conversational contexts, the current findings can be overturned by follow-up work controlling linguistic information through other means. But if not, there may be something specially informative about the combined prosodic and lexical cues to questionhood that boosts children's anticipations before they can use these cues separately. Even in adults, Torreira and colleagues (\citeyear{torreira2015}) were able to show that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural (spliced) speech. The present findings are the first to show evidence of a similar effect developmentally.

@@

The question effect itself has implications for our current theories about turn prediction. While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies \citealp{ford1996, duncan1972}. Our results rather suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are the results of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal. In other words, our results suggest that predictions are driven by what is \textit{beyond} the end of the current turn---that questions, not lexical cues, are sufficient for prediction at the first level of conversation monitoring.

>>>>

Interestingly, adults and children both were strongly affected by transition type, in that they made more anticipatory switches after hearing questions, compared to non-questions. Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}). Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}. The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to upcoming responses.

@@

However, these results do not support the idea that lexical information is sufficient for children, as has been proposed for adults previously (\citealp{de-ruiter2006}) and replicated in the present study. On the contrary, children showed significant gains in the \textit{normal} speech condition, but \textit{not} in the \textit{words only} condition. Notably, the \textit{normal} speech condition, in addition to having both lexical and prosodic information, is also the one most likely to occur in our participants' daily lives (compared to muffled or robotic speech) and therefore may have an extra advantage of familiarity over the other conditions.

@@

Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \citep{magyari2012, de-ruiter2006} and replicated in the current study for adult participants: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal. Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues. If this effect arose in our data because children do not make predictions based on phonetically manipulated speech in conversational contexts, the current findings can be overturned by follow-up work controlling linguistic information through other means. But if not, there may be something specially informative about the combined prosodic and lexical cues to questionhood that boosts children's anticipations before they can use these cues separately. Even in adults, Torreira and colleagues (\citeyear{torreira2015}) were able to show that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural (spliced) speech. The present findings are the first to show evidence of a similar effect developmentally.

@@

The question effect itself has implications for our current theories about turn prediction. While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies (\citealp{ford1996, duncan1972}). Our results rather suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are the results of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal. In other words, our results suggest that predictions are driven by what is \textit{beyond} the end of the current turn---that questions, not lexical cues, are sufficient for prediction at the first level of conversation monitoring.
